Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-04-04 PAPER.tex
revised text about N_bls and N_days --- moved a sqrt(2) factor around

where $\nu$ are frequencies in GHz (\citealt{thompson_et_al2001}). We use a receiver temperature of $144$\,K, yielding $T_{\rm sys} = 431$\,K at $150$\,MHz. 
This is lower than the $T_{\rm sys}$ of $500$\,K used in \citetalias{ali_et_al2015} because of several small mis-calculation errors that were 
identified\footnote{For example, there was a missing a square root in going from a variance to a standard deviation.}.
\item $\sqrt{2}$: This factor in the denominator of the sensitivity equation comes from taking the real part of the power spectrum 
estimates after cross-multiplying independent ``even" and ``odd" visibility measurements (this cross-multiplication is done principally to avoid a noise bias). In \citetalias{ali_et_al2015}, a factor of $2$ was mistakenly used.
\item $N_{\rm lst}$: The number of independent LST bins that go into a power spectrum estimation. The sensitivity scales as the square root 
because we integrate incoherently over time. For PAPER-64, $N_{\rm lst} = 8$.
\item $N_{\rm seps}$: The number of baseline separation types (where baselines of a unique separation type have the same orientation and length) averaged incoherently in a final power spectrum estimate. For the 
analysis in this paper, we only use one type of baseline (PAPER's 30\,m East/West baselines), hence $N_{\rm seps}=1$. The updated limits in Kolopanis et al. (\textit{submitted})
use three separation types.
\item $t_{\rm int}$: Length of an independent integration of the data. It is crucial to adapt this number if filtering is applied along the time axis (i.e. a 
fringe-rate filter). We compute the effective integration time of our fringe-rate filtered data by scaling the original integration time $t_{i}$
using the following:
\begin{equation}
t_{\rm int} = t_{i} \frac{\int1 \, df}{\int w^{2}(f) \,df},
\end{equation}
where $t_{i}=43$ seconds, $t_{\rm int}$ is the fringe-rate filtered integration time, $w$ is the fringe-rate profile, and the integral is 
taken over all fringe-rates. For PAPER-64, this number is $t_{\rm int} = 3857$\,s. 
\item $N_{\rm days}$: The total number of days of data analyzed. In \citetalias{ali_et_al2015}, this number was set to $135$. However, because we 
divide our data in half (to form ``even" and ``odd" datasets), this number should be reduced by a factor of $2$. Additionally, 
because our LST coverage is not $100\%$ complete (it doesn't overlap for every single day), we compute a more realistic value of 
$N_{\rm days}$ as:
\begin{equation}
 \frac{1}{N_{\rm days}} = \sqrt{\Big\langle\frac{1}{N_{i}^{2}} \Big\rangle_{i}},
 \end{equation}
\noindent where $i$ indexes LST (\citealt{jacobs_et_al2015}). For PAPER-64, our revised estimate of $N_{\rm days}$ is $\sim34$ 
days.
\item $N_{\rm bls}$: The number of baselines contributing to the sensitivity of a power spectrum estimate. In \citetalias{ali_et_al2015}, this number was 
the total number of $30$\,m East/West baselines used in the analysis. However, using the total number of baselines ($N_{\rm bls\_total}$) neglects 
the fact that we average baselines into groups for computational speed-up when cross-multiplying data. Our revised estimate for the parameter is:
\begin{equation}
N_{\rm bls} = \frac{N_{\rm bls\_total}}{N_{\rm gps}}\sqrt{N_{\rm gps}^{2}-N_{\rm gps}},
\end{equation}
\noindent where, as with the \citetalias{ali_et_al2015} analysis, $N_{\rm gps} = 5$. Each baseline group averages down linearly as the number of baselines 
entering the group ($N_{\rm bls\_total}/N_{\rm gps}$) and then as the square root of the number of cross-multiplied pairs ($\sqrt{N_{\rm gps}^{2} - 
N_{\rm gps}}$). For the revised PAPER-64 analysis with only one baseline separation type, this becomes $N_{\rm bls} \sim 46$ instead 
of $51$. 
\item $N_{\rm pols}$: The number of polarizations averaged together. For the case of Stokes I, $N_{\rm pols}=2$.
\end{itemize}

@@

\subsubsection{Mitigating Bias}

>>>>

where $\nu$ are frequencies in GHz (\citealt{thompson_et_al2001}). We use a receiver temperature of $144$\,K, yielding $T_{\rm sys} = 431$\,K at $150$\,MHz. 
This is lower than the $T_{\rm sys}$ of $500$\,K used in \citetalias{ali_et_al2015} because of several small mis-calculation errors that were 
identified\footnote{For example, there was a missing a square root in going from a variance to a standard deviation.}.
\item $\sqrt{2}$: This factor in the denominator of the sensitivity equation comes from taking the real part of the power spectrum 
estimates after cross-multiplying independent ``even" and ``odd" visibility measurements (this cross-multiplication is done principally to avoid a noise bias). In \citetalias{ali_et_al2015}, a factor of $2$ was mistakenly used.
\item $N_{\rm lst}$: The number of independent LST bins that go into a power spectrum estimation. The sensitivity scales as the square root 
because we integrate incoherently over time. For PAPER-64, $N_{\rm lst} = 8$.
\item $N_{\rm seps}$: The number of baseline separation types (where baselines of a unique separation type have the same orientation and length) averaged incoherently in a final power spectrum estimate. For the 
analysis in this paper, we only use one type of baseline (PAPER's 30\,m East/West baselines), hence $N_{\rm seps}=1$. The updated limits in Kolopanis et al. (\textit{submitted})
use three separation types.
\item $t_{\rm int}$: Length of an independent integration of the data. It is crucial to adapt this number if filtering is applied along the time axis (i.e. a 
fringe-rate filter). We compute the effective integration time of our fringe-rate filtered data by scaling the original integration time $t_{i}$
using the following:
\begin{equation}
t_{\rm int} = t_{i} \frac{\int1 \, df}{\int w^{2}(f) \,df},
\end{equation}
where $t_{i}=43$ seconds, $t_{\rm int}$ is the fringe-rate filtered integration time, $w$ is the fringe-rate profile, and the integral is 
taken over all fringe-rates. For PAPER-64, this number is $t_{\rm int} = 3857$\,s. 
\item $N_{\rm days}$: The total number of days of data analyzed. In \citetalias{ali_et_al2015}, this number was set to $135$. However, because we 
divide our data in half (to form ``even" and ``odd" datasets, or $N_{\rm datasets} = 2$), this number should reflect the number of days in each individual dataset instead of the total. Additionally, this number should be adjusted to reflect the actual number of cross-multiplications that occur between datasets (``even" with ``odd" and ``odd" with ``even", but not ``odd" with ``odd" or ``even" with ``even", for reasons explained in Section \ref{sec:MitBias}). Finally, because our LST coverage is not $100\%$ complete (it doesn't overlap for every single day), we incorporate a root-mean-square statistic in computing a realistic value of 
$N_{\rm days}$. Our expression therefore becomes:
\begin{equation}
N_{\rm days} = \sqrt{\langle N_{i}^{2}\rangle} \sqrt{(N_{\rm datasets}^{2}-N_{\rm datasets})}
 %\frac{1}{N_{\rm days}} = \sqrt{\Big\langle\frac{1}{N_{i}^{2}} \Big\rangle_{i}},
 \end{equation}
\noindent where $i$ indexes LST and frequency channel over all datasets (\citealt{jacobs_et_al2015}). For PAPER-64, our revised estimate of $N_{\rm days}$ is $\sim47$ 
days.
\item $N_{\rm bls}$: The number of baselines contributing to the sensitivity of a power spectrum estimate. In \citetalias{ali_et_al2015}, this number was 
the total number of $30$\,m East/West baselines used in the analysis. However, using the total number of baselines ($N_{\rm bls\_total}$) neglects 
the fact that we average baselines into groups for computational speed-up when cross-multiplying data. Our revised estimate for the parameter is:
\begin{equation}
N_{\rm bls} = \frac{N_{\rm bls\_total}}{N_{\rm gps}}\sqrt{\frac{N_{\rm gps}^{2}-N_{\rm gps}}{2}},
\end{equation}
\noindent where, as with the \citetalias{ali_et_al2015} analysis, $N_{\rm gps} = 5$. Each baseline group averages down linearly as the number of baselines 
entering the group ($N_{\rm bls\_total}/N_{\rm gps}$) and then as the square root of the number of cross-multiplied pairs \Big($\sqrt{\frac{N_{\rm gps}^{2} - 
N_{\rm gps}}{2}}$\Big). For the revised PAPER-64 analysis with only one baseline separation type, this becomes $N_{\rm bls} \sim 32$ instead 
of $51$. 
\item $N_{\rm pols}$: The number of polarizations averaged together. For the case of Stokes I, $N_{\rm pols}=2$.
\end{itemize}

@@

\subsubsection{Mitigating Bias}
\label{sec:MitBias}
